---
title: "Project 3: PartA Tutorial"
author: "Yiying Gao"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{PartA Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

This package is the first part for UW STAT 302 Project 3. In this package, it 
includes four functions, my_t.test, my_lm, my_knn_cv and my_rf_cv, and also two data, my_penguins and my_gapminder.

To use this package, installment is the required step, which could be done by installing from GitHub. 

The below code could be used to install from GitHub.
```{r, eval = FALSE, message = FALSE}
devtools::install_github("viggieG/PartA")
```

To begin, let's load our package and other related ones used in package functions.
```{r, eval = FALSE, message = FALSE}
library(PartA)
library(class)
library(randomForest)
library(dplyr)
library(ggplot2)
library(tidyr)
```

After finish all the steps above, package function tutorial now begins!

## Tutorial for my_t.test

For this part, it illustrates three conditions for my_t.test function, two.sided, greater and less. At the same time, we set the p-value cut-off to be α=0.05.

The below code demonstrates the alternative hypothesis that mu is not equal to 60
```{r, message = FALSE}
library(PartA)
my_t_test(my_gapminder$lifeExp, "two.sided", 60)
```

For this condition, since the p_val is greater than 0.05, we do not reject the null hypothesis.

The below code demonstrates the alternative hypothesis that mu is less than 60
```{r, message = FALSE}
library(PartA)
my_t_test(my_gapminder$lifeExp, "less", 60)
```

For this condition, since the p_val is less than 0.05, we reject the null hypothesis.

The below code demonstrates the alternative hypothesis that mu is greater than 60
```{r, message = FALSE}
library(PartA)
my_t_test(my_gapminder$lifeExp, "greater", 60)
```

For this condition, since the p_val is greater than 0.05, we do not reject the null hypothesis.

## Tutorial for my_lm
```{r, message = FALSE, warning = FALSE}
library(tibble)
# construct the model
pred_model <- my_lm(lifeExp ~ gdpPercap + continent, my_gapminder)
pred_model
```

From the above results, we can see that gdpPercap predictor variables are statistically significant because its p-values are all less than 0.05. 
All of its results' signs show positive, so it is positively related to the response variable.

For this linear model, the p-value for each independent variable tests two null hypothesis. One is that the variable has no correlation with the dependent variable. It is the same for gdpPercap. This null hypothesis for it is that the gdpPercap has no correlation with the lifeExp, but the p value of it is less than 0.05. Thus gdpPercap sample data provide enough evidence to reject the null hypothesis for the entire population, and show there is a non-zero correlation between them.

The second one is whether or not the gdpPercap coefficient is equal to zero. Considering its p value is less than 0.05, so gdpPercap sample data provide enough evidence to reject the null. That is, the gdpPercap coefficient is not equal to zero.

```{r, message = FALSE}
library(ggplot2)
# construct the fitted model data
mod_fits <- model.matrix(lifeExp ~ gdpPercap + continent, my_gapminder) %*%               pred_model[, 1] 

# store the actual and fitted data
my_df <- data.frame(actual = my_gapminder$lifeExp, fitted = mod_fits)

# draw the actual vs. fitted graph
ggplot(my_df, aes(x = fitted, y = actual)) +
  geom_point() +
  # draw the trend line
  geom_abline(slope = 1, intercept = 0, col = "red", lty = 2) + 
  theme_bw(base_size = 15) +
  # draw the title and axes
  labs(x = "Fitted values", y = "Actual values", title = "Actual vs. Fitted") +
  theme(plot.title = element_text(hjust = 0.5))
```

From the graph above, we can clearly see that the model fitting situation is not happy. Multiple actual values match one single fitted value at the same time. Though this model indeed makes correct predictions at some values, data dispersion is too big to make up.

## Tutorial for my_knn_cv

Before getting into this part, I would like to introduce what the process of cross-validation is doing and why it is useful. 

According to its specific concept, cross-validation is a resampling procedure used to evaluate machine learning models on a limited data sample. The procedure has a single parameter called k that refers to the number of groups that a given data sample is to be split into. It will repeat k times. Each time it will use 1/k group data as test data and remaining as the training data to build the model.

The reason why it is useful is that it gives model the opportunity to train on multiple train-test splits. This gives a better indication of how well model will perform on unseen data. Hold-out, on the other hand, is dependent on just one train-test split.

That's the end for introduction. Now let's begin our trip!

```{r, message = FALSE, warning = FALSE}
library(tidyverse)

# clean all NAs from the data
my_penguins <- drop_na(my_penguins)

# select the covariats we need
train_penguins <- my_penguins %>%
  select(bill_length_mm,bill_depth_mm,flipper_length_mm,body_mass_g)

# construct the model with different k_nn
knn1 <- my_knn_cv(train_penguins, my_penguins$species, 1, 5)
knn2 <- my_knn_cv(train_penguins, my_penguins$species, 2, 5)
knn3 <- my_knn_cv(train_penguins, my_penguins$species, 3, 5)
knn4 <- my_knn_cv(train_penguins, my_penguins$species, 4, 5)
knn5 <- my_knn_cv(train_penguins, my_penguins$species, 5, 5)
knn6 <- my_knn_cv(train_penguins, my_penguins$species, 6, 5)
knn7 <- my_knn_cv(train_penguins, my_penguins$species, 7, 5)
knn8 <- my_knn_cv(train_penguins, my_penguins$species, 8, 5)
knn9 <- my_knn_cv(train_penguins, my_penguins$species, 9, 5)
knn10 <- my_knn_cv(train_penguins, my_penguins$species, 10, 5)

# build a matrix to show the training
TM <- c(mean(knn1[1] != my_penguins$species), mean(knn2[1] != my_penguins$species), mean(knn3[1] != my_penguins$species), mean(knn4[1] != my_penguins$species), mean(knn5[1] != my_penguins$species), mean(knn6[1] != my_penguins$species), mean(knn7[1] != my_penguins$species), mean(knn8[1] != my_penguins$species), mean(knn9[1] != my_penguins$species), mean(knn10[1] != my_penguins$species))

# build a matrix to show the CV
CV <- c(knn1[2], knn2[2], knn3[2], knn4[2], knn5[2], knn6[2], knn7[2], knn8[2], knn9[2], knn10[2])
CV

```

Based on the training misclassification rates, I would choose the model with 
k_nn = 10, because it has the lowest taining rates, which is caused by the high 
sensitive for the actual values. Based on the CV misclassification rates, I 
would choose the model with k_nn = 1, because it has the lowest CV rates.

I would choose the model with k_nn = 1 in practice, because it has the lowest CV rates, and CV rates can more clearly show the the percentage of classifications that were incorrect comparing to the real data than the training ones.

## Tutorial for my_rf_cv
```{r, message = FALSE, warning = FALSE}
# construct three list to store the results
pred1 <- 0
pred2 <- 0
pred3 <- 0

# run each k thirty times and store
for (i in 1 : 30) {
  pred1 <- append(pred1, my_rf_cv(2))
  pred2 <- append(pred2, my_rf_cv(5))
  pred3 <- append(pred3, my_rf_cv(10))
}

# clean the 0 from the list
pred1 <- pred1[-1]
pred2 <- pred2[-1]
pred3 <- pred3[-1]

# create the number of simulations used in graphing
ob1 <- rep(1 : 30, each = 2)
ob2 <- rep(1 : 30, each = 5)
ob3 <- rep(1 : 30, each = 10)

# create each k's data frame
k2 <- data_frame("observation" = ob1,
                      "CV" = pred1)
k5 <- data_frame("observation" = ob2,
                      "CV" = pred2)
k10 <- data_frame("observation" = ob3,
                      "CV" = pred3)

# draw the box plot for k = 2
ggplot(data = k2, 
       aes(x = observation, y = CV )) +
  geom_boxplot(fill = "lightblue") +
  theme_bw(base_size = 20) +
  labs(title = "CV Estimated of K = 2", 
       x = "Simulations", 
       y = "MSE") +
  theme(plot.title =
          element_text(hjust = 0.5))

# draw the box plot for k = 5
ggplot(data = k5, 
       aes(x = observation, y = CV )) +
  geom_boxplot(fill = "lightblue") +
  theme_bw(base_size = 20) +
  labs(title = "CV Estimated of K = 5", 
       x = "Simulations", 
       y = "MSE") +
  theme(plot.title =
          element_text(hjust = 0.5))

# draw the box plot for k = 10
ggplot(data = k10, 
       aes(x = observation, y = CV )) +
  geom_boxplot(fill = "lightblue") +
  theme_bw(base_size = 20) +
  labs(title = "CV Estimated of K = 10", 
       x = "Simulations", 
       y = "MSE") +
  theme(plot.title =
          element_text(hjust = 0.5))

# build a matrix to store the mean and std deviation
mean_std <- matrix(c(mean(pred1), sd(pred1), mean(pred2), sd(pred2), mean(pred3), sd(pred3)),ncol = 2, nrow = 3)

# classify the data meaning and k
colnames(mean_std) <- c("Mean","Std dev")
rownames(mean_std) <- c("k = 2", "k = 5", "k = 10")
mean_std
```

From the above boxplots and the table, we can clearly see that with increasing 
k, the mean of MSE is decreasing, like k = 2 and k = 10. Also, with the 
increasing k, the standard deviation is increasing. This situation is caused by
the bias and variance tradeoff. With higher k, Our bias is decreasing because we 
are training our model to be more specific to our training data. The variance is
increasing because our model will be overfit, and will vary substantially with 
new training data.

The k = 5 shows different moves probably because its data has high dispersion 
and lower number of extreme values. For it, it does show much lower mean and 
extremely high standard deviation. However, if simply compare k = 5 to k = 2, it
still follows the above rules: higher k, lower mean and higher standard deviation.




